The Astropy Project: \\ Sustaining and Growing a Community-oriented Open-source Project and the Latest Major Release (v5.0) of the Core Package

### 2021-12-02 src/ms.tex
title attempt
some minor tweaks

\title{The Astropy Project: xx}

@@

The \python programming language is a high-level, interpreted (versus compiled)
programming language that has become an industry standard across many
computational domains, technological sectors, and fields of research.
Despite claims to the contrary \citep{Portegies-Zwart:2020}, \python enables
scalable, time- and energy-efficient code execution \citep[e.g.,][]{Augier:2021}
with a focus on code readability, ease of use, and interoperability with other
languages.
Over the last decade, \python has grown enormously in popularity to become a
dominant programming language in the astronomical and broader scientific
communities.
For example, Figure~\ref{fig:python-mentions} shows the number of yearly
full-text mentions of \python as compared to a few other programming languages
in refereed articles in the astronomical literature, demonstrating its nearly
exponential growth in popularity.
The rapid adoption of \python by astronomy researchers, students, observatories,
and technical staff combined with an associated increase in awareness and
interest about open-source software tools is contributing to a paradigm shift in
the way research is done, data is analyzed, and results are shared in astronomy
and beyond.

One of the factors that has led to its rapid ascent in popularity in scientific
contexts has been the significant, volunteer-driven effort behind developing
community-oriented open-source software tools and fostering communities of users
and developers that have grown around these efforts.
Today, a broad and feature-diverse ``ecosystem'' of packages exists in the
\python scientific computing landscape: Roughly ordered from general-use to
domain-specific, this landscape now includes packages that provide core
numerical analysis functionality like \package{numpy} \citep{todo} and
\package{scipy} \citep{todo}, to visualization frameworks like
\package{matplotlib} \citep{todo}, to machine learning and data analysis
packages like \package{tensorflow} \citep{todo}, \package{pymc3} \citep{todo},
and \package{emcee} \citep{todo}, to specialized, domain-specific packages like
\package{yt} \citep{todo}, \package{plasmapy} \citep{todo}, \package{sunpy}
\citep{todo}, \package{Biopython} \citep{todo}, and \package{sympy}
\citep{sympy} (to name a few in each category).
The \astropypkg \citep{astropy:2013, astropy:2018} core package began as an
effort to consolidate the development of commonly-used functionality needed to
perform astronomical research into a community-developed \python package.

>>>>

\title{The Astropy Project: \\
       Sustaining and Growing a Community-oriented Open-source Project and
       the Latest Major Release (v5.0) of the Core Package}

@@

The \python programming language is a high-level, interpreted (as opposed to
compiled) programming language that has become an industry standard across many
computational domains, technological sectors, and fields of research.
Despite claims to the contrary \citep{Portegies-Zwart:2020}, \python enables
scalable, time- and energy-efficient code execution \citep[e.g.,][]{Augier:2021}
with a focus on code readability, ease of use, and interoperability with other
languages.
Over the last decade, \python has grown enormously in popularity to become a
dominant programming language in the astronomical and broader scientific
communities.
For example, Figure~\ref{fig:python-mentions} shows the number of yearly
full-text mentions of \python as compared to a few other programming languages
in refereed articles in the astronomical literature, demonstrating its nearly
exponential growth in popularity.
The rapid adoption of \python by astronomy researchers, students, observatories,
and technical staff combined with an associated increase in awareness and
interest about open-source software tools is contributing to a paradigm shift in
the way research is done, data is analyzed, and results are shared in astronomy
and beyond.

One of the factors that has led to its rapid ascent in popularity in scientific
contexts has been the significant, volunteer-driven effort behind developing
community-oriented open-source software tools and fostering communities of users
and developers that have grown around these efforts.
Today, a broad and feature-diverse ``ecosystem'' of packages exists in the
\python scientific computing landscape: Roughly ordered from general-use to
domain-specific, this landscape now includes packages that provide core
numerical analysis functionality like \package{numpy} \citep{todo} and
\package{scipy} \citep{todo}, visualization frameworks like \package{matplotlib}
\citep{todo}, machine learning and data analysis packages like
\package{tensorflow} \citep{todo}, \package{pymc3} \citep{todo}, and
\package{emcee} \citep{todo}, domain-specific libraries like \package{yt}
\citep{todo}, \package{plasmapy} \citep{todo}, \package{sunpy} \citep{todo},
\package{Biopython} \citep{todo}, and \package{sympy} \citep{sympy} (to name a
few in each category).
The \astropypkg \citep{astropy:2013, astropy:2018} core package began in this
vein, as an effort to consolidate the development of commonly-used functionality
needed to perform astronomical research into a community-developed \python
package.
